---
title: "Separation stats"
author: "Sasha and Flavia"
output:
  html_document:
    df_print: paged
---

# Plot data

```{r}
ggplot(statsumm, aes(treat, e.rel.130, color = trial)) + 
  geom_point() + 
  geom_line(aes(group = trial)) + 
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```

Through means.

```{r}
ggplot(statsumm, aes(treat, e.rel.130, color = trial)) + 
  geom_line() + 
  geom_point() + 
  stat_summary(aes(group = trial), fun = mean, geom = 'line') +
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```


```{r}
ggplot(statsumm, aes(as.integer(factor(treat)), e.rel.130, color = trial)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```

```{r}
ggplot(statsumm, aes(treat, e.rel.130, color = trial)) + 
  geom_point() + 
  geom_line() + 
  geom_smooth(se = FALSE) +
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```


# Mixed-effects models

## Model 1
First ignore any variation in separation response among trials (plot shows evidence that we should *not* ignore it).

```{r}
mod1 <- lmer(e.rel.130 ~ separation * slurry.type + (1|trial), data = statsumm, REML = FALSE)
summary(mod1)
anova(mod1)
```

```{r}
plot(mod1)
qqnorm(resid(mod1))
qqline(resid(mod1))
```

## Model 2
Now include this random separation effect across trials.

```{r}
mod2 <- lmer(e.rel.130 ~ separation * slurry.type + (1 + separation | trial), data = statsumm, REML = FALSE)
summary(mod2)
anova(mod2)
mod2

```

## Model 2 hypothsis tests

We can determine if there is evidence for it with a likelihood ratio test.

```{r}
anova(mod1, mod2, test = 'Chisq')
```

Strong evidence, with p value below 1E-6.

For interaction term:

```{r}
mod2b <- lmer(e.rel.130 ~ separation + slurry.type + (1 + separation | trial), data = statsumm, REML = FALSE)
anova(mod2, mod2b, test = 'Chisq')
```

p < 0.01, pretty clear.

And it is clear from the plots that there is no consistent difference between

* solid and liquid fractions from AD,
* raw and solid from pig slurry.

The only thing not clear from the combination of these results and the plots is if.

* AD liquid fraction is in fact lower than raw AD, 
* pig liquid fraction is in fact lower than raw pig slurry. 

We can test exactly these and only these by grouping factor levels and using likelihood ratio tests.


```{r}
statsumm[, separation.psg := separation]
statsumm[slurry.type == 'PS' & separation == 'Liquid', separation.psg := 'Unsep']
table(statsumm[, .(slurry.type, separation, trial)])
table(statsumm[, .(slurry.type, separation.psg, trial)])
```

```{r}

statsumm[, separation.adg := separation]
statsumm[slurry.type == 'AD' & separation == 'Liquid', separation.adg := 'Unsep']
table(statsumm[, .(slurry.type, separation, trial)])
table(statsumm[, .(slurry.type, separation.adg, trial)])

```

```{r}
mod2c <- lmer(e.rel.130 ~ separation.psg + slurry.type + (1 + separation.psg | trial), data = statsumm, REML = FALSE)
mod2d <- lmer(e.rel.130 ~ separation.adg + slurry.type + (1 + separation.adg | trial), data = statsumm, REML = FALSE)
anova(mod2, mod2c, test = 'Chisq')
anova(mod2, mod2d, test = 'Chisq')
```

Singular fit is problematic for comparison.

```{r}
mod1c <- lmer(e.rel.130 ~ separation.psg + slurry.type + (1 | trial), data = statsumm, REML = FALSE)
mod1d <- lmer(e.rel.130 ~ separation.adg + slurry.type + (1 | trial), data = statsumm, REML = FALSE)
anova(mod1, mod1c, test = 'Chisq')
anova(mod1, mod1d, test = 'Chisq')
```

No better.
Try subset instead.

```{r}
statsummul <- statsumm[separation %in% c('Unsep', 'Liquid')]
```

```{r}
mod3 <- lmer(e.rel.130 ~ separation * slurry.type + (1|trial), data = statsummul, REML = FALSE)
summary(mod3)
anova(mod3)
```

Check for interaction now.

```{r}
mod3b <- lmer(e.rel.130 ~ separation + slurry.type + (1|trial), data = statsummul, REML = FALSE)
anova(mod3, mod3b, test = 'Chisq')
```

Nope!
So solid fraction response drives interaction.
That is obvious from the plots.

So we can easily test for overall unsep vs. liquid effect.

```{r}
mod3c <- lmer(e.rel.130 ~ slurry.type + (1|trial), data = statsummul, REML = FALSE)
summary(mod3c)
anova(mod3c, mod3b, test = 'Chisq')
```

We still have the singular fit issue, but a very clear effect.

# Conclusions:
* Separation affects emission
* The reduction for the solid fraction differs between slurry types
* The liquid fraction has lower emission than whole slurry for both, and the reduction is not clearly different between slurry types
* For digestate, both liquid and solid fractions have lower emission than whole slurry, but are not clearly different from each other, if we consider that the separation effect varies among trials (and it clearly does, from mod2)
* For pig slurry, whole slurry and the solid fraction are not clearly different, but liquid is clearly different (lower) than both




#Stats for the biochar trials
Make the plots of liquid pig slurry and liquid digestate with an increasing biochar dose

```{r}
ggplot(bsumm, aes(dose, e.rel.130, color = trial)) + 
  geom_point() + 
  geom_line(aes(group = trial)) + 
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```
```{r}
ggplot(bsumm, aes(dose, e.rel.130, color = trial)) + 
  geom_line() + 
  geom_point() + 
  stat_summary(aes(group = trial), fun = mean, geom = 'line') +
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```
```{r}
ggplot(bsumm, aes(as.numeric(dose), e.rel.130, color = trial)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```
```{r}
ggplot(bsumm, aes(dose, e.rel.130, color = trial)) + 
  geom_point() + 
  geom_line() + 
  geom_smooth(se = FALSE) +
  facet_wrap(~ slurry.type, scales = 'free_x') + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```
Linear model to test if the increasing dose of biochar (0, 2, 10 %) added to liquid digestate and liquid pig slurry has an effect on emissions. 
```{r}
modb <- lm(e.rel.130 ~dose:slurry.type, data = bsumm)
summary(modb)
anova(modb)
```

from the model and plots it looks like biochar addition had an effect only on liquid digestate only

```{r}
plot(modb)
qqnorm(resid(modb))
qqline(resid(modb))
```

#test separately liquid pig and liquid digestate to make sure it is true

```{r}

#subset pig slurry data and write the linear model
bsumm.pig <- bsumm[trial %in% paste('Trial', 9), ]
modb2 <- lm(e.rel.130 ~ dose, data = bsumm.pig)
summary(modb2)
anova(modb2)

```
#no effect of digestate dose on liquid pig slurry

```{r}
#subset digestate slurry data and write the linear model
bsumm.ad <- bsumm[trial %in% paste('Trial', 8), ]
modb3 <- lm(e.rel.130 ~ dose, data = bsumm.ad)
summary(modb3)
anova(modb3)
```
There is an effect of biochar dose on liquid digestate

Conclusions
*For the liquid fraction of digestate, the addition of an increasing dose of biochar had an effect on emissions
*An increasing dose of biochar increased the emissions from the liquid fraction of digestate
*An increasing dose of biochar to the liquid fraction of pig slurry did not have an effect on emissions



```{r}
# Plot
plot(bsumm.ad$dose, bsumm.ad$e.rel.130,
     xlab = "Biochar Dose (% wet weight)",
     ylab = "losse (% TAN)",
     main = "",
     pch = 16,
     col = "blue",
     cex = 1)

# Add regression line
abline(modb3, col = "red", lwd = 2)
```
```{r}
ggplot(bsumm.solid, aes(dose, e.rel.130, color =dose )) + 
  geom_point() + 
  geom_line(aes(group = dose)) + 
  theme_bw() + 
  labs(y = 'Loss (frac. of TAN)') + 
  theme(legend.title = element_blank())
```
from this plot it looks like biochar addition caused higher emissions compared to non-addition



Fit the linear model to test if the emissions are influenced by biochar addition

```{r}
modb4 <- lm(e.rel.130 ~ dose, data = bsumm.solid)
summary(modb4)
anova(modb4)
```
#Conclusions
*Addition of biochar to the solid fraction of separated digestate had an effect on emissions
*Solid fraction of digestate with biochar addition had higher emissions compared to solid fraction of digestate without biochar addition
